A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management

### 2022-12-03 papers/earths-future/submission_03.tex
all done?

However, as discussed in \cref{sec:introduction}, estimates of trade-offs between desired performance metrics (\eg, cost and reliability) depend on probabilistic models of uncertainty.
In this paper we present a method for integrating the \glspl{sow} used in exploratory modeling into a formal decision analytic framework using a subjective probability distribution over the space of possible futures, which is used to infer implicit weights over the \glspl{sow}.
This approach is particularly suited for problems where the \glspl{sow} are generated from or conditioned on specific scenarios (\eg, \gls{rcp} scenarios) or where there are multiple models of the underlying processes (\eg, multiple models of the response of local sea levels to global temperature), which can lead to the ``multiple \gls{pdf} problem'' (discussed in \cref{sec:analysis-condition}).
A motivating advantage is that it makes assumptions about the likelihoods of different \glspl{sow} transparent to decision-makers.

@@

One way to interpret an ensemble of \glspl{sow} is as {iid} draws from some probabilistic data generating process.
This commonly arises when a single deep uncertainty (\eg, an emissions pathway) is used as an input for a stochastic model.
To clarify language, we draw a distinction between a \gls{sow}, which is a single realization of a possible future, and a \emph{scenario}, which is a model that can be used to generate \glspl{sow}.
These terms are defined relative to the problem at hand; for example, a global emissions trajectory may be a \gls{sow} from an integrated assessment model, but when it is used as a deterministic input to an ice sheet model we would describe it as a scenario.

@@

We can also apply this theoretical lens to examine the approach, common in \gls{dmdu} applications, of sampling parameters from a set of fixed ranges.
The \emph{scenario} in this case is thus the choice of bounds on the parameters; it is consistent with our above definition of a scenario because it provides a probabilistic model that can be used to sample \glspl{sow}.
For example, \citeA{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
Similarly, \citeA{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range.
Analyses that use this methodology are implicitly assuming a single probabilistic model in which different variables are drawn from independent Uniform distributions.
Many limitations of Uniform and other noninformative priors have been documented in the literature, including (i) that they can induce unrealistic implicit priors over functions of parameters and (ii) results are sensitive to the parameterization of a given process \cite{seaman_noninformative:2012}.
Yet while replacing Uniform distributions with alternatives such as maximum-entropy distributions can address some of these challenges \cite<\eg,>{gupta_prior:2022}, subjective modeling choices remain necessary.
Our primary concern here is not that these subjective modeling assumptions are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation \cite<see>[regarding the importance of iterative critique]{gelman_workflow:2020} they may lead to inscrutable decision processes and poor decisions.

@@

We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citeA{ruckert_coastal:2019}.
The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \cite{wong_brick0.2:2017}, (iii) the \citeA{kopp_probabilistic:2014} model (``K14''), and (iv) the \citeA{deconto_antarctica:2016} model (``DP16'').
The \citeA{kopp_probabilistic:2014} and \citeA{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \cite{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying probabilistic scenarios of \gls{msl}.

@@

The choices of physical model and \gls{rcp} scenario jointly determine future \gls{msl} $p(\overline{y}|t)$.
\Cref{fig:lsl-evolution}(a) shows the time-varying 90\% credible intervals of \gls{msl} for three representative models.
The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen probabilistic scenarios considered.
The stark differences between different scenarios of \gls{slr} arise primarily from different representations of Antarctic Ice Sheet contributions to global \gls{slr} and statistical calibration methodologies.
For a more detailed discussion we refer the reader to \citeA{ruckert_coastal:2019}.
We return in \cref{sec:results-conditional} to the challenge of decision making given multiple \glspl{pdf}.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{tradeoffs-by-rcp}
  \caption{
    Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
    For emphasis, we highlight three representative models: the Brick Slow model \protect\cite{wong_brick0.2:2017} under \gls{rcp} 2.6, the K14 \protect\cite{kopp_probabilistic:2014} model under \gls{rcp} 6.0 and the DP16 model \protect\cite{deconto_antarctica:2016,kopp_evolving:2017} under \gls{rcp} 8.5.
    (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs.
    (B): trade-off between up-front cost and lifetime expected damages (eq.~\ref{eq:led}).
    Light gray lines show estimates for all 16 models (four \gls{rcp} scenarios times four physical parameterizations) considered.
    Colored lines highlight three representative models for emphasis.
    The gray arrows indicate the direction of preference.
  }\label{fig:tradeoffs-by-rcp}
\end{figure}

@@

This approach is, in a sense, another form of exploratory modeling: instead of considering a very large ensemble of \glspl{sow}, we consider a much smaller set of probabilistic models.
Scenario-conditional analysis can be attractive because it allows modelers to focus on their domain expertise (\eg, the response of ice sheets and global sea levels to a particular climate future).
However, conditioning simulations on a set of climate futures and physical models presents what we term ``the multiple \gls{pdf} problem'' because it leaves decision makers with many \glspl{pdf} to choose from and hence many trade-off curves to navigate.
The multiple \gls{pdf} problem has also been discussed in other contexts.
For example, \citeA{sharma_rcp:2021} model the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
Similarly, \citeA{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios. As a last example, \citeA{haasnoot_sealevelrise:2021} identify global adaptation needs for different \gls{slr} scenarios.
Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}
Since not all house owners or contractors have expertise in assessing the relative likelihood of different climate futures, they may not be well positioned to make this decision.

>>>>

However, as discussed in \cref{sec:introduction}, estimates of trade-offs between desired performance metrics (\eg, cost and reliability) depend on probabilistic models of uncertainty.
In this paper we present a method for integrating the \glspl{sow} used in exploratory modeling into a formal decision analytic framework using a subjective probability distribution over the space of possible futures, which is used to infer implicit weights over the \glspl{sow}.
This approach is particularly suited for problems where the \glspl{sow} are generated from or conditioned on specific scenarios (\eg, \gls{rcp} scenarios) or where there are multiple models of the underlying processes (\eg, multiple parameterizations of the response of local sea levels to global temperature), which can lead to the ``multiple \gls{pdf} problem'' (discussed in \cref{sec:analysis-condition}).
A motivating advantage is that it makes assumptions about the likelihoods of different \glspl{sow} transparent to decision-makers.

@@

One way to interpret an ensemble of \glspl{sow} is as {iid} draws from some probabilistic data generating process.
This commonly arises when a single deep uncertainty (\eg, an emissions pathway) is used as an input for a stochastic model.
To clarify language, we draw a distinction between a \emph{\gls{sow}}, which is a single realization of a possible future, and a \emph{probabilistic scenario}, which we define loosely as a set of assumptions for which probabilistic projections (\ie, \glspl{sow}) are available or can be simulated.
For example, in the case study described in \cref{sec:case-slr} we use four physical models of the processes relevant to \gls{slr} and four \gls{rcp} scenarios to generate a total of 16 probabilistic scenarios.

@@

We can also apply this theoretical lens to examine the approach, common in \gls{dmdu} applications, of sampling parameters from a set of fixed ranges.
The \emph{scenario} in this case is thus the choice of bounds on the parameters; it is consistent with our above definition of a probabilistic scenario because \glspl{sow} can be sampled using probabilistic methods.
For example, \citeA{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
Similarly, \citeA{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range.
Analyses that use this methodology are implicitly assuming a single probabilistic model in which different variables are drawn from independent Uniform distributions.
Many limitations of Uniform and other noninformative priors have been documented in the literature, including (i) that they can induce unrealistic implicit priors over functions of parameters and (ii) results are sensitive to the parameterization of a given process \cite{seaman_noninformative:2012}.
Yet while replacing Uniform distributions with alternatives such as maximum-entropy distributions can address some of these challenges \cite<\eg,>{gupta_prior:2022}, subjective modeling choices remain necessary.
Our primary concern here is not that these subjective modeling assumptions are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation \cite<see>[regarding the importance of iterative critique]{gelman_workflow:2020} they may lead to inscrutable decision processes and poor decisions.

@@

We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citeA{ruckert_coastal:2019}.
The four physical process models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \cite{wong_brick0.2:2017}, (iii) the \citeA{kopp_probabilistic:2014} model (``K14''), and (iv) the \citeA{deconto_antarctica:2016} model (``DP16'').
The \citeA{kopp_probabilistic:2014} and \citeA{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \cite{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying probabilistic scenarios of \gls{msl}.

@@

The choices of physical model and \gls{rcp} scenario jointly determine future \gls{msl} $p(\overline{y}|t)$.
\Cref{fig:lsl-evolution}(a) shows the time-varying 90\% credible intervals of \gls{msl} for three representative models.
The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen probabilistic scenarios considered.
The stark differences between different scenarios of \gls{slr} arise primarily from different representations of Antarctic Ice Sheet contributions to global \gls{slr} and statistical calibration methodologies.
For a more detailed discussion we refer the reader to \citeA{ruckert_coastal:2019}.
We return in \cref{sec:results-conditional} to the challenge of decision making given multiple scenarios.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{tradeoffs-by-rcp}
  \caption{
    Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
    For emphasis, we highlight three representative models: the Brick Slow model \protect\cite{wong_brick0.2:2017} under \gls{rcp} 2.6, the K14 \protect\cite{kopp_probabilistic:2014} model under \gls{rcp} 6.0 and the DP16 model \protect\cite{deconto_antarctica:2016,kopp_evolving:2017} under \gls{rcp} 8.5.
    (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs.
    (B): trade-off between up-front cost and lifetime expected damages (eq.~\ref{eq:led}).
    Light gray lines show estimates for all 16 models (four \gls{rcp} scenarios $\times$ four physical process models) considered.
    Colored lines highlight three representative models for emphasis.
    The gray arrows indicate the direction of preference.
  }\label{fig:tradeoffs-by-rcp}
\end{figure}

@@

This approach is, in a sense, another form of exploratory modeling: instead of considering a very large ensemble of \glspl{sow}, we consider a much smaller set of probabilistic scenarios.
Scenario-conditional analysis can be attractive because it allows modelers to focus on their domain expertise (\eg, the response of ice sheets and global sea levels to a particular climate future).
However, conditioning simulations on a set of climate futures and physical models presents what we term ``the multiple \gls{pdf} problem'' because it leaves decision makers with many \glspl{pdf} to choose from and hence many trade-off curves to navigate.
The multiple \gls{pdf} problem has also been discussed in other contexts.
For example, \citeA{sharma_rcp:2021} model the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
Similarly, \citeA{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios. As a last example, \citeA{haasnoot_sealevelrise:2021} identify global adaptation needs for different \gls{slr} scenarios.
Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which probabilistic scenario to design for onto the end user.}
Since not all house owners or contractors have expertise in assessing the relative likelihood of different climate futures, they may not be well positioned to make this decision.
